---
title: "Deliverable 3 - Trophic Level Analysis"
author: "Niko, Reina, and Braylee"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
library(here)
library(tidyverse)
```

[KF has now done this!] need to adjust collectData for df5 and combine collect year and collect month
```{r}
# Step 1: Read the CSV files into separate data frames
df1 <- read_csv(here("data", "processed", "alg_biomass_means.csv"))
df2 <- read_csv(here("data", "processed", "final_fish_count_means.csv"))
df3 <- read_csv(here("data", "processed", "waterqual_means.csv"))
df4 <- read_csv(here("data", "processed", "SW_Nitrate_means.csv"))
df5 <- read_csv(here("data", "processed", "nutrient_data_means.csv"))
```

```{r}
# Step 2: Combine the data frames into one
merged_data_all <- full_join(df1, df2) %>% 
  full_join(., df3) %>% 
  full_join(., df4) %>% 
  full_join(., df5)

# Check the first few rows of the merged data frame
head(merged_data_all)

# Step 3: Write the merged data frame into a new CSV file
write_csv(merged_data_all, here("data", "processed", "merged_data.csv"))

```


**KF: Unclear why you set the model up this way (with mean ash free dry mass as the predictor variable and surface water nitrate concentrations as the response)... I think you want to reverse your variables!**

**Also be sure to include plots/tests that check that data are meeting the assumptions of linear models!**
```{r}

linear_model1 <- lm(mean_SWN ~ mean_adjAshFreeDryMass, data = merged_data_all)
summary(linear_model1)
```

A linear regression was performed to predict the relationship between nitrate concentration and algal biomass. The results of the linear regression indicated that nitrate concentration explained -11.94% of the variation in algal biomass (F(1,6)=0.2535, p=0.6326, R^2=-0.05871).



```{r}
linear_model2 <- lm(mean_weight ~ mean_adjAshFreeDryMass, data = merged_data_all)
summary(linear_model2)
```

A linear regression was performed to predict the relationship between fish abundance and algal biomass. The results of the linear regression indicated that algal biomass explained 24.96% of the variation in fish abundance (F(1,9)=4.327, p=0.06725, R^2=0.2202). 

**KF: Here and below, check that the predictor variables in the regression are the same variables you're putting on your x-axes!**
```{r}
ggplot(data = merged_data_all,
       aes(x = mean_weight, 
           y = mean_adjAshFreeDryMass)) +
  labs(x= "Mean Fish Biomass (g)",
       y = "Mean Algal Biomass (g)")+
  geom_point() +
  theme_classic() +
  geom_smooth(method = "lm")
```
Figure 1. The affect of Mean Fish Biomass on Mean Algal Biomass across 3 sites(CRAM, TOOK, PRLA) from 2016-2019. The trend line indicates that as Fish Biomass increases, Algal Biomass decreases with the grey section representing a 95% confidence interval to the line of best fit. (F(1,9)=4.37, p=0.0657, R^2=0.250) 

```{r}
ggplot(data = merged_data_all,
       aes(x = mean_SWN, 
           y = mean_adjAshFreeDryMass)) +
  labs(x= "Surface Water Nitrate (micromoles per Liter)", 
       y = "Mean Algal Biomass (g)")+
  geom_point() +
  theme_classic() +
  geom_smooth(method = "lm")
```
Figure 2. The effect of Mean Surface Water Nitrate on Mean Algal Biomass across 3 sites (CRAM, TOOK, PRLA) from 2016-2019. The trend line indicates that as Surface Water Nitrogen levels increase, Algal Biomass increases as well with many variations in the data as Nitrate concentrations increase (95% confidence interval). the higher  (F(1,6)=0.254, p=0.633, R^2= -0.119)
```{r}
par(mfrow = c(1, 3))
plot(linear_model1, which = c (1:3))
```

```{r}
par(mfrow = c(1, 3))
plot(linear_model2, which = c (1:3))
```

```{r}
linear_model1 <- lm(total_individuals ~ mean_adjAshFreeDryMass, data = merged_data_all)
summary(linear_model1)
```
```{r}
ggplot(data = merged_data_all,
       aes(x = total_individuals, 
           y = mean_adjAshFreeDryMass)) +
  labs(x= "Total Fish Abundance ",
       y = "Mean Algal Biomass (g)")+
  geom_point() +
  theme_classic() +
  geom_smooth(method = "lm")
```

